Uncertainty propagation in deep neural networks for sparse coding

### 2018-03-09 paper/BayesianLISTA.tex
conclusion reviewed

\section{Discussion}
\label{sec:discussion}
As far as authors aware, this is the first implementation of Bayesian deep sparse coding algorithm. Although there are works on Bayesian sparsity in context of neural networks \citep{he2017bayesian}, they are not the Bayesian neural networks in the same sense as Bayesian \textsc{lista} but rather the interpretation of the sparse Bayesian  learning algorithm as the long short-term memory network. We find not only correct predictions but also useful posterior estimates for the predictions that show how the model is confident in its decision.

The inference is based on expectation propagation algorithm and though it is not suited for distributed inference, stochastic variant of assumed density filtering can potentially be used \citep{li2015stochastic}. This would allow the proposed approach to scale.

The original \textsc{lista} algorithm can be used for inverse sparse problems with both overcomplete dictionaries as in the original paper, or undercomplete dictionaries \citep{borgerding2017amp}. Though this paper presents only results with undercomplete dictionaries, this is not the limitation of the method and it can be used for overcomplete dictionaries in a similar manner.

\section{Conclusions and future work}
\label{sec:conclusions}
We have presented the new method for propagating the uncertainty through the soft thresholding function. To achieve this goal we have approximated the outputs of the soft thresholding function with spike and slab distribution, and we have shown that this distribution can stay within the same family after linear transformation with Gaussian weights and inputs. This allowed us to propose the Bayesian \textsc{lista} algorithm that learns the distributions of the weights and parameters and makes the uncertainty estimates of the outputs. The forward propagation in the algorithm is based on the proposed uncertainty propagation method, the backward propagation is based on the probabilistic backpropagation method, that was modified to account for likelihood of the Bayesian \textsc{lista} and its recurrent nature.

Experiments on the synthetic and \textsc{mnist} datasets demonstrate that in addition to the posterior estimates the algorithm can achieve better quality of predictions than the original \textsc{lista} when the available data is small. Experiments on active learning demonstrate that the posterior estimates by the algorithm are adequate as the uncertainty estimated with it allows to more optimally select training points.

In future we plan to estimate of shrinkage parameter $\lambda$ within the backpropagation with ADF, use minibatches for learning and use stochastic expectation propagation for distributed inference. 
\bibliography{bibliography}
\bibliographystyle{icml2018}

>>>>

\section{Discussion}
\label{sec:discussion}
As far as the authors aware, this is the first implementation of Bayesian deep sparse coding algorithm. Although there are works on Bayesian sparsity in context of neural networks \citep{he2017bayesian}, they are not the Bayesian neural networks in the same sense as Bayesian \textsc{lista} but rather the interpretation of the sparse Bayesian  learning algorithm as the long short-term memory network. We find not only correct predictions but also useful posterior estimates for the predictions that show how the model is confident in its decisions.

%The inference is based on the expectation propagation algorithm and though it is not suited for distributed inference, stochastic variant of assumed density filtering can potentially be used \citep{li2015stochastic}. This would allow the proposed approach to scale.

The \textsc{lista}-based algorithms can be applied for the sparse coding problem with both overcomplete dictionaries as in the original paper \citep{gregor2010learning}, and undercomplete dictionaries \citep{borgerding2017amp}. Though this paper presents only results with undercomplete dictionaries, this is not the limitation of the proposed method and it can be used for overcomplete dictionaries in a similar manner.

\section{Conclusions and future work}
\label{sec:conclusions}
We have presented the new method for propagating the uncertainty through the soft thresholding function. %To achieve this goal 
We have approximated the outputs of the function with a spike and slab distribution, and we have shown that this distribution can stay within the same family after linear transformation with Gaussian weights and inputs of a neural network. This allowed us to propose the Bayesian \textsc{lista} network and efficient inference algorithm that learns the distributions of the weights and makes the uncertainty estimates of the outputs. The forward propagation in the algorithm is based on the proposed uncertainty propagation method, the backward propagation is based on the probabilistic backpropagation method, that was remarkably expanded to account for multidimensionality of inputs and outputs, likelihood of the Bayesian \textsc{lista} and its recurrent nature.

Experiments on the synthetic and \textsc{mnist} datasets demonstrate that the proposed algorithm preserves the predictive accuracy of non-Bayesian methods while also providing posterior estimates. We also show that when the training data is very small the proposed algorithm significantly outperforms the classical \textsc{lista} in terms of predictive accuracy. Experiments on active learning demonstrate that the proposed Baysian \textsc{lista} gives accurate posterior estimates that can be used for selection of a next data point where a label should be obtain for.

The shrinkage parameter $\lambda$ is currently treated as a deterministic hyperparameter for the Baysian \textsc{lista}. In future we plan to incorporate it inside the model treating it as a random variable. For this we need to extend both the uncertainty propagation method to include uncertainty of $\lambda$ and the probabilistic backpropagation algorithm to estimate its posterior. We will study stochastic expectation propagation applicability in order to replace ADF that may improve further the quality of posterior estimates~\citep{li2015stochastic}.    
\bibliography{bibliography}
\bibliographystyle{icml2018}
